Analysis of NLA17 Dissolved Gas Indicator

### 2021-05-21 scripts/dgIndicatorAnalysis.Rmd
reconciling merge conflict

dg %>% 
  ggplot(aes(as.factor(x = WSA9), y = log(dissolved.n2o.nmol))) + geom_boxplot()

dg %>% 
  ggplot(aes(as.factor(x = WSA9), y = (n2o.sat.ratio))) + geom_boxplot()
```

@@

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, WSA9, WSA9_NAME)

names(dg_1)
dg_1 <- clean_names(dg_1) %>% filter(wsa9 != "AK") %>% 
  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
# underscores are "safer" names used in GIS

@@

# Check CRS
st_crs(ecoR)
names(ecoR)
# changed from 4326 to 5070
ecoR1 <- st_transform(ecoR, 5070) %>% 
  select(OBJECTID, WSA9, WSA9_NAME) # Grab variables of interest
st_crs(ecoR1)

@@

## Summarizing distances
Getting distances between sites overall and by ecoregion to help set distances for lagged scatterplots.  First, Worked out steps on a sample of NLA sites.  Could only get parts of Anna Springsteen's code to work.
```{r, eval=FALSE, echo=FALSE, distances}
names(dg.sf)
head(dg.sf$WSA9_NAME)
# distinct returns all unique geometry + attribute combinations.  convert to df if not interested in geometry
dg.sf %>% as.data.frame() %>% distinct(WSA9_NAME) 
# distinct.Spatial(dg.sf, WSA9_NAME) # this doesn't work either.
distinct(dg, WSA9_NAME)

@@

names(ecoR)
class(ecoR)
# contrast ecoRname <- as.data.frame(select(ecoR,"WSA9_NAME"))
# geom is sticky with select but not pull
# ecoRname <- as.data.frame(pull(ecoR,"WSA9_NAME"))
ecoRname1 <- as.character(pull(ecoR,"WSA9_NAME")) #Springsteen code needs to be character to populate columns of matrix below
class(ecoRname1)
head(ecoRname1, 9)
# initialize summary tables
euc.summary <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
dim(euc.summary)
colnames(euc.summary) <- c("ALL",ecoRname1) #indicates same assignment in each object
rownames(euc.summary) <- c("Min","Q1","Median","Q3","Max")
euc.summary

@@

count(dg.sf, WSA9_NAME) %>% print(n = Inf)

# Get ecoregion-specific NLA sites
cpl <- dg.sf %>% filter(WSA9_NAME == "Coastal Plains")

@@

# nap # nap object doesn't exist
nap <- dg.sf %>% filter(WSA9_NAME == "Northern Appalachians")
nap_sp <- as(nap, "Spatial")
class(nap_sp)
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

@@

# npl 
npl <- dg.sf %>% filter(WSA9_NAME == "Northern Plains")
npl_sp <- as(npl, "Spatial")
class(npl_sp)
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

@@

# sap
sap <- dg.sf %>% filter(WSA9_NAME == "Southern Appalachians")
sap_sp <- as(sap, "Spatial")
class(sap_sp)
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

@@

# spl
spl <- dg.sf %>% filter(WSA9_NAME == "Southern Plains")
spl_sp <- as(spl, "Spatial")
class(spl_sp)
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

@@

# tpl
tpl <- dg.sf %>% filter(WSA9_NAME == "Temperate Plains")
tpl_sp <- as(tpl, "Spatial")
class(tpl_sp)
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

@@

# umw
umw <- dg.sf %>% filter(WSA9_NAME == "Upper Midwest")
umw_sp <- as(umw, "Spatial")
class(umw_sp)
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))

@@

# wmt
wmt <- dg.sf %>% filter(WSA9_NAME == "Western Mountains")
wmt_sp <- as(wmt, "Spatial")
class(wmt_sp)
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

@@

# xer
xer <- dg.sf %>% filter(WSA9_NAME == "Xeric")
xer_sp <- as(xer, "Spatial")
class(xer_sp)
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
# 10 km bin/lag only has 2 points plotted so start with 50 km
hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

@@

The code below is to be run separately for each ecoregion.  The CPL, TPL, and UMW ecoregions show strong positive spatial autocorrelation.  Negative spatial autocorrelation occurs in the NPL ecoregion because of the maximum outlier at a South Dakota site.  The SPL and XER ecoregions show no evidence of spatial autocorrelation.  Interestingly, it appears that the ecoregions with the highest dissolved.n2o concentrations, CPL, TPL, and UMW, drive the national pattern.
```{r, knn_moran}
names(dg)
# also includ WSA9, WSA9_NAME feature
dg_1 <- select(dg,site.id:dissolved.n2o.nmol, WSA9, WSA9_NAME)
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
distinct(dg_1, wsa9)
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

@@

st_crs(dg_sf1)
glimpse(dg_sf1)
# Get ecoregion-specific NLA sites
## CPL
cpl <- dg_sf1 %>% filter(wsa9 == "CPL" & visit_no == 1 & sample_source == "DG")

@@

## NAP
nap <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1 & sample_source == "DG")

@@

## NPL
npl <- dg_sf1 %>% filter(wsa9 == "NPL" & visit_no == 1 & sample_source == "DG")

@@

## SAP
sap <- dg_sf1 %>% filter(wsa9 == "SAP" & visit_no == 1 & sample_source == "DG")

@@

## SPL
spl <- dg_sf1 %>% filter(wsa9 == "SPL" & visit_no == 1 & sample_source == "DG")

@@

## TPL
tpl <- dg_sf1 %>% filter(wsa9 == "TPL" & visit_no == 1 & sample_source == "DG")

@@

## UMW
umw <- dg_sf1 %>% filter(wsa9 == "UMW" & visit_no == 1 & sample_source == "DG")

@@

## WMT
wmt <- dg_sf1 %>% filter(wsa9 == "WMT" & visit_no == 1 & sample_source == "DG")

@@

## XER
xer <- dg_sf1 %>% filter(wsa9 == "XER" & visit_no == 1 & sample_source == "DG")

@@

st_crs(dg_sf1)
names(dg_sf1)
nap_visit1 <- dg_sf1 %>% filter(wsa9 == "NAP" & visit_no == 1)
class(nap_visit1)
st_crs(nap_visit1)

@@

# trying ncf package which uses resampling to specify number of null permutations under the null to assess significance levels.  See if significance persist after increas resamp from 100 to 1000, which is the default.
# The x-intercept in the results is the distance at which objects are no more similar than expected by chance alone across the region.
# Run on CPL as that is giving correlation greater than 1, which according to Fletcher and Fortin might be due to small number of pairs of points used in that first distance bin or to outliers.  Perhaps, cluster of high Delaware sites (n visit 1 = 5) causes the correlation to be greater than 1.
# Looks like 71 pairs of points being used.
# With a lag, or bin distance, of 50 km only go out 10 bins, 500 km, and do a Bonferroni adjustment of 0.05/10 = 0.005
names(dg_1)
dg_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1)
# N.B. default is resamp = 1000 so that can take 10 minutes for n =1090 sites
class(dg_cplvisit1)

@@

# CPL outlier check by working through all 10 pairwise removal of cluster of 5 high Delaware sites
out_cplvisit1 <- dg_1 %>% filter(wsa9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))

@@

# trying pgirmess package uses library spdep including moran.test and geary.test functions.  Distances are euclidean and in the same units as the spatial coordinates.  As run below, using geographical coordinates in decimal degrees so are distance classes retuned in decimal degrees?
# what if run on a spatial points data frame?
# Uses normality assumption, which is faster, to assess significance
dg_napvisit1 <- dg_1 %>% filter(wsa9 == "NAP" & visit_no == 1)
coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
correl_napn2o_pg <- pgirmess::correlog(coords_nap, log(dg_napvisit1$dissolved_n2o_nmol), method = "Moran", nbclass = NULL, alternative = "two.sided")

@@

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, WSA9, WSA9_NAME, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3))
## Subset dg_1 by visit_no == 1
dg_visit1 <- dg_1 %>% dplyr::filter(visit_no == 1)

## Perform the same test for each wsa9 using Z = "log_n2o"
## Start by defining a function
ncf_correlog <- function(dat, # data to send to ncf::correlog
                         z_col, # name of the column in the dataset for water quality parameter
                         i = 50, # increment to be used in ncf::correlog
                         N = 1000, # number of resamples in ncf::correlog
                         x_col = "map_lon_dd", # name of the column in the dataset for longitude
                         y_col = "map_lat_dd", # name of the column in the dataset for latitude
                         lat_lon = TRUE, # if the x and y coordinates for ncf::correlog are lat and lon
                         rem_missing = FALSE, # argument of ncf::correlog
                         run_quietly = FALSE){ # argument of ncf::correlog
  library(ncf)

@@

## Split the dataframe by each wsa9: turns the data frame into a list 
##   where each list item is a data frame filtered to that wsa9
## lapply applies the function ncf_correlog on each list object (a data frame)
## the final product is a list of correlog objects
## this is similar to looping
## setting seed so get same result from permutations ncf::correlog function
set.seed(2035)
log_n2o_correlog_list <- dg_visit1 %>%
  split(.$wsa9) %>% # '.' means use the data frame that preceded this line
  lapply(function(x) ncf_correlog(dat = x, z_col = "log_n2o")) 
#ncf_correlog is the function name, 
#dat is the paramater that takes a data.frame,
#z_col is the parameter that takes a string name for a column of a water quality paramater

@@

## combine into one dataframe to grab the wsa9 (notice the name no longer has an s at the end)

log_n2o_correlog_dat <- log_n2o_correlog_dats %>%
  dplyr::bind_rows(., .id = "wsa9")

str(log_n2o_correlog_dat)
head(log_n2o_correlog_dat)
## Make individual plots, the following object is a list containing a ggplot for each
log_n2o_correlog_plots <- log_n2o_correlog_dat %>%
  split(.$wsa9) %>% # splits the data frame by wsa9, which creates a list with 9 data frames
  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each wsa9 to make a plot
                                   plot_title = NULL,
                                   n_col = "n", 
                                   mean_col = "mean",
                                   cor_col = "cor",
                                   xint_col = "xint",
                                   pval_col = "pval",
                                   name_col = "wsa9",
                                   title_size = 14, 
                                   subtitle_size = 14, 
                                   axis_title_size = 14, 
                                   axis_text_size = 14))
## if hard to see writing run again but increase "title_size", "subtitle_size", "axis_title_size", "axis_text_size" as needed

@@

## alternatively, can make a facet plot since log_n20_correlog_dat is a dataframe in the correct format
names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "wsa9"

## create column to shape point, sig == TRUE means two-sided test < 0.05
log_n2o_correlog_dat$sig <- ifelse(log_n2o_correlog_dat$pval * 2 < 0.05,
                                   TRUE, FALSE) 
head(log_n2o_correlog_dat)
## can change wsa9 to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
# log_n20_correlog_dat$wsa9 <- factor(log_n20_correlog_dat, 
#                                     levels = c("CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER"))

## make facet plot
log_n2o_correlog_facetplot <- ggplot(log_n2o_correlog_dat) +
  geom_hline(yintercept = 0, color = "gray", linetype = "solid") +
  # geom_point(aes(x=mean, y=cor, shape=sig), size=I(2), color="black") +
  # geom_line(aes(x=mean, y=cor, group=1), size=I(0.5), color="black", linetype="solid") +
  lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1), 
                        size=I(2), color = "black", 
                        linetype = "solid", linesize = I(0.5), 
                        linecolor = "black") +
  scale_shape_manual(values=c("TRUE"=19, "FALSE"=1), 
                     labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
                     guide=FALSE) +
  facet_wrap(~wsa9, ncol=3, scales="free") +
  labs(x="Distance (mean-of-class)", y="Correlation") +
  theme_bw() +
  theme(plot.title = element_text(size=14, hjust=0.5, 
                                  face="bold", color="black"),
        plot.subtitle = element_text(size=14, hjust=0.5, color="black"),
        axis.title = element_text(size=14, face="plain", color="black"),
        axis.text = element_text(size=14, color="black"),
        strip.text = element_text(size=14, color="black", face="bold"),
        strip.background = element_rect(fill="gray", color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
print(log_n2o_correlog_facetplot)
## can save the plot using 
# cowplot::ggsave(filename = "FILENAME2.png", #saves a png version of the figure to the working folder
#                 plot = log_n20_correlog_facetplot) 
## can change .png to other forms
## can play with scale, width, and height to adjust the output of the figure saved to the png file

@@

The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
# also includ WSA9, WSA9_NAME feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, wsa9, wsa9_name, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & wsa9 != "AK")

>>>>

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = log(dissolved.n2o.nmol))) + geom_boxplot()

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()
```

@@

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm)

names(dg_1)
dg_1 <- clean_names(dg_1) %>% filter(ag.eco9.nm != "AK") %>% 
  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
# underscores are "safer" names used in GIS

@@

# Check CRS
st_crs(ecoR)
names(ecoR)
# changed from 4326 to 5070
ecoR1 <- st_transform(ecoR, 5070) %>% 
  select(OBJECTID, ag.eco9.nm, ag.eco9.nm) # Grab variables of interest
st_crs(ecoR1)

@@

## Summarizing distances
Getting distances between sites overall and by ecoregion to help set distances for lagged scatterplots.  First, Worked out steps on a sample of NLA sites.  Could only get parts of Anna Springsteen's code to work.
```{r, eval=FALSE, echo=FALSE, distances}
names(dg.sf)
head(dg.sf$ag.eco9.nm)
# distinct returns all unique geometry + attribute combinations.  convert to df if not interested in geometry
dg.sf %>% as.data.frame() %>% distinct(ag.eco9.nm) 
# distinct.Spatial(dg.sf, ag.eco9.nm) # this doesn't work either.
distinct(dg, ag.eco9.nm)

@@

names(ecoR)
class(ecoR)
# contrast ecoRname <- as.data.frame(select(ecoR,"ag.eco9.nm"))
# geom is sticky with select but not pull
# ecoRname <- as.data.frame(pull(ecoR,"ag.eco9.nm"))
ecoRname1 <- as.character(pull(ecoR,"ag.eco9.nm")) #Springsteen code needs to be character to populate columns of matrix below
class(ecoRname1)
head(ecoRname1, 9)
# initialize summary tables
euc.summary <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
dim(euc.summary)
colnames(euc.summary) <- c("ALL",ecoRname1) #indicates same assignment in each object
rownames(euc.summary) <- c("Min","Q1","Median","Q3","Max")
euc.summary

@@

count(dg.sf, ag.eco9.nm) %>% print(n = Inf)

# Get ecoregion-specific NLA sites
cpl <- dg.sf %>% filter(ag.eco9.nm == "Coastal Plains")

@@

# nap # nap object doesn't exist
nap <- dg.sf %>% filter(ag.eco9.nm == "Northern Appalachians")
nap_sp <- as(nap, "Spatial")
class(nap_sp)
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

@@

# npl 
npl <- dg.sf %>% filter(ag.eco9.nm == "Northern Plains")
npl_sp <- as(npl, "Spatial")
class(npl_sp)
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

@@

# sap
sap <- dg.sf %>% filter(ag.eco9.nm == "Southern Appalachians")
sap_sp <- as(sap, "Spatial")
class(sap_sp)
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

@@

# spl
spl <- dg.sf %>% filter(ag.eco9.nm == "Southern Plains")
spl_sp <- as(spl, "Spatial")
class(spl_sp)
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

@@

# tpl
tpl <- dg.sf %>% filter(ag.eco9.nm == "Temperate Plains")
tpl_sp <- as(tpl, "Spatial")
class(tpl_sp)
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

@@

# umw
umw <- dg.sf %>% filter(ag.eco9.nm == "Upper Midwest")
umw_sp <- as(umw, "Spatial")
class(umw_sp)
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))

@@

# wmt
wmt <- dg.sf %>% filter(ag.eco9.nm == "Western Mountains")
wmt_sp <- as(wmt, "Spatial")
class(wmt_sp)
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

@@

# xer
xer <- dg.sf %>% filter(ag.eco9.nm == "Xeric")
xer_sp <- as(xer, "Spatial")
class(xer_sp)
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
# 10 km bin/lag only has 2 points plotted so start with 50 km
hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

@@

The code below is to be run separately for each ecoregion.  The CPL, TPL, and UMW ecoregions show strong positive spatial autocorrelation.  Negative spatial autocorrelation occurs in the NPL ecoregion because of the maximum outlier at a South Dakota site.  The SPL and XER ecoregions show no evidence of spatial autocorrelation.  Interestingly, it appears that the ecoregions with the highest dissolved.n2o concentrations, CPL, TPL, and UMW, drive the national pattern.
```{r, knn_moran}
names(dg)
# also includ ag.eco9.nm, ag.eco9.nm feature
dg_1 <- dg
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
distinct(dg_1, ag_eco9_nm)
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

@@

st_crs(dg_sf1)
glimpse(dg_sf1)
# Get ecoregion-specific NLA sites
## CPL
cpl <- dg_sf1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & sample_source == "DG")

@@

## NAP
nap <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1 & sample_source == "DG")

@@

## NPL
npl <- dg_sf1 %>% filter(ag_eco9_nm == "NPL" & visit_no == 1 & sample_source == "DG")

@@

## SAP
sap <- dg_sf1 %>% filter(ag_eco9_nm == "SAP" & visit_no == 1 & sample_source == "DG")

@@

## SPL
spl <- dg_sf1 %>% filter(ag_eco9_nm == "SPL" & visit_no == 1 & sample_source == "DG")

@@

## TPL
tpl <- dg_sf1 %>% filter(ag_eco9_nm == "TPL" & visit_no == 1 & sample_source == "DG")

@@

## UMW
umw <- dg_sf1 %>% filter(ag_eco9_nm == "UMW" & visit_no == 1 & sample_source == "DG")

@@

## WMT
wmt <- dg_sf1 %>% filter(ag_eco9_nm == "WMT" & visit_no == 1 & sample_source == "DG")

@@

## XER
xer <- dg_sf1 %>% filter(ag_eco9_nm == "XER" & visit_no == 1 & sample_source == "DG")

@@

st_crs(dg_sf1)
names(dg_sf1)
nap_visit1 <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1)
class(nap_visit1)
st_crs(nap_visit1)

@@

# trying ncf package which uses resampling to specify number of null permutations under the null to assess significance levels.  See if significance persist after increas resamp from 100 to 1000, which is the default.
# The x-intercept in the results is the distance at which objects are no more similar than expected by chance alone across the region.
# Run on CPL as that is giving correlation greater than 1, which according to Fletcher and Fortin might be due to small number of pairs of points used in that first distance bin or to outliers.  Perhaps, cluster of high Delaware sites (n visit 1 = 5) causes the correlation to be greater than 1.
# Looks like 71 pairs of points being used.
# With a lag, or bin distance, of 50 km only go out 10 bins, 500 km, and do a Bonferroni adjustment of 0.05/10 = 0.005
names(dg_1)
dg_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1)
# N.B. default is resamp = 1000 so that can take 10 minutes for n =1090 sites
class(dg_cplvisit1)

@@

# CPL outlier check by working through all 10 pairwise removal of cluster of 5 high Delaware sites
out_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))

@@

# trying pgirmess package uses library spdep including moran.test and geary.test functions.  Distances are euclidean and in the same units as the spatial coordinates.  As run below, using geographical coordinates in decimal degrees so are distance classes retuned in decimal degrees?
# what if run on a spatial points data frame?
# Uses normality assumption, which is faster, to assess significance
dg_napvisit1 <- dg_1 %>% filter(ag_eco9 == "NAP" & visit_no == 1)
coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
correl_napn2o_pg <- pgirmess::correlog(coords_nap, log(dg_napvisit1$dissolved_n2o_nmol), method = "Moran", nbclass = NULL, alternative = "two.sided")

@@

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3))
## Subset dg_1 by visit_no == 1
dg_visit1 <- dg_1 %>% dplyr::filter(visit_no == 1)

## Perform the same test for each ag.eco9.nm using Z = "log_n2o"
## Start by defining a function
ncf_correlog <- function(dat, # data to send to ncf::correlog
                         z_col, # name of the column in the dataset for water quality parameter
                         i = 50, # increment to be used in ncf::correlog
                         N = 1000, # number of resamples in ncf::correlog
                         x_col = "map_lon_dd", # name of the column in the dataset for longitude
                         y_col = "map_lat_dd", # name of the column in the dataset for latitude
                         lat_lon = TRUE, # if the x and y coordinates for ncf::correlog are lat and lon
                         rem_missing = FALSE, # argument of ncf::correlog
                         run_quietly = FALSE){ # argument of ncf::correlog
  library(ncf)

@@

## Split the dataframe by each ag.eco9.nm: turns the data frame into a list 
##   where each list item is a data frame filtered to that ag.eco9.nm
## lapply applies the function ncf_correlog on each list object (a data frame)
## the final product is a list of correlog objects
## this is similar to looping
## setting seed so get same result from permutations ncf::correlog function
set.seed(2035)
log_n2o_correlog_list <- dg_visit1 %>%
  split(.$ag_eco9_nm) %>% # '.' means use the data frame that preceded this line
  lapply(function(x) ncf_correlog(dat = x, z_col = "log_n2o")) 
#ncf_correlog is the function name, 
#dat is the paramater that takes a data.frame,
#z_col is the parameter that takes a string name for a column of a water quality paramater

@@

## combine into one dataframe to grab the ag.eco9.nm (notice the name no longer has an s at the end)

log_n2o_correlog_dat <- log_n2o_correlog_dats %>%
  dplyr::bind_rows(., .id = "ag.eco9.nm")

str(log_n2o_correlog_dat)
head(log_n2o_correlog_dat)
## Make individual plots, the following object is a list containing a ggplot for each
log_n2o_correlog_plots <- log_n2o_correlog_dat %>%
  split(.$ag.eco9.nm) %>% # splits the data frame by ag.eco9.nm, which creates a list with 9 data frames
  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each ag.eco9.nm to make a plot
                                   plot_title = NULL,
                                   n_col = "n", 
                                   mean_col = "mean",
                                   cor_col = "cor",
                                   xint_col = "xint",
                                   pval_col = "pval",
                                   name_col = "ag.eco9.nm",
                                   title_size = 14, 
                                   subtitle_size = 14, 
                                   axis_title_size = 14, 
                                   axis_text_size = 14))
## if hard to see writing run again but increase "title_size", "subtitle_size", "axis_title_size", "axis_text_size" as needed

@@

## alternatively, can make a facet plot since log_n20_correlog_dat is a dataframe in the correct format
names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "ag.eco9.nm"

## create column to shape point, sig == TRUE means two-sided test < 0.05
log_n2o_correlog_dat$sig <- ifelse(log_n2o_correlog_dat$pval * 2 < 0.05,
                                   TRUE, FALSE) 
head(log_n2o_correlog_dat)
## can change ag.eco9.nm to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
# log_n20_correlog_dat$ag.eco9.nm <- factor(log_n20_correlog_dat, 
#                                     levels = c("CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER"))

## make facet plot
log_n2o_correlog_facetplot <- ggplot(log_n2o_correlog_dat) +
  geom_hline(yintercept = 0, color = "gray", linetype = "solid") +
  # geom_point(aes(x=mean, y=cor, shape=sig), size=I(2), color="black") +
  # geom_line(aes(x=mean, y=cor, group=1), size=I(0.5), color="black", linetype="solid") +
  lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1), 
                        size=I(2), color = "black", 
                        linetype = "solid", linesize = I(0.5), 
                        linecolor = "black") +
  scale_shape_manual(values=c("TRUE"=19, "FALSE"=1), 
                     labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
                     guide=FALSE) +
  facet_wrap(~ag.eco9.nm, ncol=3, scales="free") +
  labs(x="Distance (mean-of-class)", y="Correlation") +
  theme_bw() +
  theme(plot.title = element_text(size=14, hjust=0.5, 
                                  face="bold", color="black"),
        plot.subtitle = element_text(size=14, hjust=0.5, color="black"),
        axis.title = element_text(size=14, face="plain", color="black"),
        axis.text = element_text(size=14, color="black"),
        strip.text = element_text(size=14, color="black", face="bold"),
        strip.background = element_rect(fill="gray", color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
print(log_n2o_correlog_facetplot)
## can save the plot using 
# cowplot::ggsave(filename = "FILENAME2.png", #saves a png version of the figure to the working folder
#                 plot = log_n20_correlog_facetplot) 
## can change .png to other forms
## can play with scale, width, and height to adjust the output of the figure saved to the png file

@@

The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
# also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag.eco9.nm, ag.eco9.nm, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag.eco9.nm != "AK")
